---
title: "4 October Inclass"
author: "Triet Tran"
date: "March 7, 2017"
output:
  word_document: default
  html_document: default
---




# Project Goals

The goal of this project is to work through building functions around the Jackkinfe. We will start with some simple functions and then determine a pattern to build a more general function in the end. 


## Jackknife of the Mean!

- Have an estimator $\hat{\theta}$ of parameter $\theta$  
  want the standard error of our estimate, $se_{\hat{\theta}}$
- The jackknife approximation:
    + omit case $i$, get estimate $\hat{\theta}_{(-i)}$
    + Take the variance of all the $\hat{\theta}_{(-i)}$
    + multiply that variance by $\frac{(n-1)^2}{n}$ to get $\approx$ variance of $\hat{\theta}$
- then $se_{\hat{\theta}}=$ square root of that variance

**PHP 2560 Only** (Why $(n-1)^2 / n$?  Think about just getting the standard
    error of the mean)



Write a function called `mean.jackknife` that takes argument `a_vector` and returns a jackknife estimate of the standard error.


```{r, eval=FALSE}
a_vector<-c(2,3,1,2,1)
mean.jackknife <- function(a_vector) {
  estimate_v = NULL
  n=length(a_vector)
  for(i in 1:n){
    estimate_v[i]=(sum(a_vector)-a_vector[i])/(length(a_vector)-1)
  }
  variance<-var(estimate_v)
  adj_variance<-variance*((n-1)^2/n)
  return(sqrt(adj_variance))
}
mean.jackknife(a_vector)
```


#### Jackknife for the mean Example: test your code here


```{r, eval = FALSE}
some_normals <- rnorm(100,mean=7,sd=5)
mean(some_normals)
(formula_se_of_mean <- sd(some_normals)/sqrt(length(some_normals)))
all.equal(formula_se_of_mean,mean.jackknife(some_normals))
```



##Jackknife for Gamma Parameters


The following function is a way to calculate the method of moments estimators for the gamma distribution:

```{r, eval=FALSE}
gamma.est <- function(the_data) {
  m <- mean(the_data)
  v <- var(the_data)
  a <- m^2/v
  s <- v/m
  return(c(a=a,s=s))
}
gamma.est(a_vector)

vector=c(1,2,3,4,2)
a=NULL
s=NULL
for(i in 1:5){
a[i]=gamma.est(vector[-i])[1]
s[i]=gamma.est(vector[-i])[2]
}
a
s
```



###Jackknife for Gamma Parameters Function

Write a function called `gamma.jackknife` that takes argument `a_vector` and returns jackknife standard error estimates on the gamma parameters. 

```{r, eval=FALSE }
gamma.jackknife <- function(a_vector) {
  n=length(a_vector)
  variance = NULL
  adj_variance = NULL
  a=NULL
  s=NULL
  for(i in 1:5){
  a[i]=gamma.est(vector[-i])[1]
  s[i]=gamma.est(vector[-i])[2]
  }
  variance[1]<-var(a)
  variance[2]<-var(s)
  adj_variance<-variance*((n-1)^2/n)
  return(sqrt(adj_variance))
}
```



###Jackknife for Gamma Parameters Example


```{r, eval=FALSE}
input <- rgamma(1000, shape=0.1, scale=10)
gamma.est(input)
gamma.jackknife(input)
```



##Jackknife for linear regression coefficients

Write a function called `jackknife.lm` that takes arguments `df`, `formula` and `p` and returns jackknife standard error estimates on the coefficients of a linear regression model. 

```{r, eval=FALSE}
jackknife.lm <- function(df,formula,p) {
  n=nrow(df)
  p=ncol(df)
  coef_est<-matrix(nrow=n,ncol=p)
  for(i in 1:n){
  my.lm<-lm(formula,data=df[-i,])
  coef_est[i,]=coefficients(my.lm)
}
  variance<-apply(coef_est,2,var)
  adj_variance<-variance*((n-1)^2/n)
  return(sqrt(adj_variance))
}
```


###Jackknife for linear regression coefficients Example


```{r, eval=FALSE}
output <- 1.2 + 0.6*input +  rnorm(1000, 0, 2.1)
data <- data.frame(output,input)
my.lm <- lm(output~input, data=data)
coefficients(my.lm)
# "Official" standard errors
sqrt(diag(vcov(my.lm)))
jackknife.lm(df=data,formula="output~input",p=2)
```




##Refactoring the Jackknife


- Omitting one point or row is a common sub-task

- The general pattern:

```
figure out the size of the data
for each case
   omit that case
   repeat some estimation and get a vector of numbers
take variances across cases
scale up variances
take the square roots
```

- Refactor by extracting the common "omit one" operation

- Refactor by defining a general "jackknife" operation



###The Common Operation


- _Problem_: Omit one particular data point from a larger structure

- _Difficulty_: Do we need a comma in the index or not?

- _Solution_: Works for vectors, lists, 1D and 2D arrays, matrices, data frames:



## Goal: 


- Make the function select the correct dimensions
    - length for a 1d object
    - number of rows for 2d
- Write a function `omit.case` that omits a point given the data and returns the data minus that point. Make sure it can handle higher dimensions. 
    ```{r, eval=FALSE}
 omit.case <- function(the_data, omitted_point) {
   if(is.vector(the_data)) {
     the_data <- the_data[-omitted_point]
   } else {
     the_data <- the_data[-omitted_point,]
   }
   return(the_data)
   }
```

- Write a function `omit_and_est` that takes the data with an omitted point and returns whatever function your estimator does.  
    ```{r,eval=FALSE}
 omit_and_est <- function(omit) {
   estimator(omit.case(the_data, omitted_point = omit))
 }
```

```{r, eval=FALSE}
jackknife <- function(estimator, the_data) {

  if(is.vector(the_data)) {
     n <- length(the_data)
   } else {
     n <- dim(the_data)[1]
   }
  
  omit_and_est <- function(omit) {
   estimator(omit.case(the_data, omitted_point = omit))
  }
  
  jackknife.ests <- matrix(sapply(1:n, omit_and_est), ncol=n)
  
  var.of.reestimates <- apply(jackknife.ests, 1, var)
  
  jackknife.var <- ((n-1)^2/n)*var.of.reestimates
  
  jackknife.stderr <- sqrt(jackknife.var)
  
  return(jackknife.stderr)
}
```



---

###It works


```{r, eval=FALSE}
jackknife(estimator=mean,the_data=some_normals)
all.equal(jackknife(estimator=mean,the_data=some_normals),
          mean.jackknife(some_normals))
```

---



```{r, eval=FALSE}
all.equal(jackknife(estimator=gamma.est,the_data=data$input),
          gamma.jackknife(data$input))
```

---


```{r, eval=FALSE}
all.equal(jackknife(estimator=gamma.est,the_data=data$input),
          gamma.jackknife(data$input), check.names=FALSE)
```


```{r, eval=FALSE}
est.coefs <- function(the_data) {
  return(lm(output~input,data=the_data)$coefficients)
}
est.coefs(data)
all.equal(est.coefs(data), coefficients(my.lm))
```

---


```{r, eval=FALSE}
jackknife(estimator=est.coefs,the_data=data)
all.equal(jackknife(estimator=est.coefs,the_data=data),
          jackknife.lm(df=data,formula="output~input",p=2))
```


##Further Refactoring of jackknife()


The code for `jackknife()` is still a bit clunky:
- Ugly `if-else` for finding `n`
- Bit at the end for scaling variances down to standard errors

- write a function that calculates the `n` needed for the above code:

    ```{r, eval=FALSE}
data_size <- function(the_data) {
  n <- ifelse(is.vector(the_data), length(the_data), nrow(the_data))
}
    ```
- Write a function that calculate the variance of all the estimates and returns the standard error
```{r, eval=FALSE}
scale_and_sqrt_vars <- function(jackknife.ests, n) {
  jackknife.ests <- matrix(sapply(1:n, omit_and_est), ncol=n)
  var.of.reestimates <- apply(jackknife.ests, 1, var)
  jackknife.var <- ((n-1)^2/n)*var.of.reestimates
  jackknife.stderr <- sqrt(jackknife.var)
  return(jackknife.stderr)
}
```



###Now invoke those functions

```{r, eval=FALSE}
jackknife <- function(estimator,the_data) {
  n <- data_size(the_data)
  omit_and_est <- function(omit) {
    estimator(omit.case(the_data,omit))
  }
  jackknife.ests <- matrix(sapply(1:n, omit_and_est), ncol=n)
  return(scale_and_sqrt_vars(jackknife.ests,n))
}
```

